---
title: "Data Visualization I"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

## Loading data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```


## Basic Scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

It's blank! Because we haven't added any geoms.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

This code would accomplish the same thing:

```{r, eval = FALSE}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

This plot is helpful, kind of, but not as informative as we might want. So let's add another dimension: color! 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Let's continue to build on that by making the points a bit more transparent, and by adding a line:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
 
From the class webpage: 

Neat! The curve gives a sense of the relationship between variables, and the transparency shows where data are overlapping. I can’t help but notice, though, that the smooth curve is for all the data but the colors are only for the scatterplot. Turns out that this is due to where I defined the mappings. The X and Y mappings apply to the whole graphic, but color is currently geom-specific. Sometimes you want or need to do this, but for now I don’t like it. If I’m honest, I’m also having a hard time seeing everything on one plot, so I’m going to add facet based on name as well.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Cool! Let's try another graph to investigate the time of year:

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

### Learning assessment

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(tmax_f = 
           (tmax * (9/5)) + 32
         ) |> 
  mutate(tmin_f = 
           (tmin * (9/5)) + 32
         ) |> 
  ggplot(aes(x = tmin_f, y = tmax_f)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE)
```

## More plotting! 

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

Back to scatterplots: sometimes, you can only control for overlapping using transparency (alpha) so much before it still isn't sufficient. There are some other methods to help with that! 

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

## Aesthetics

There are lots of aesthetics! The help page for each individual geom will tell you which are available- some (color) apply to all the geoms, while others (shape) apply to only some geoms. 

### Learning assessment

WHy don't these two lines of code produce the same graphs? 

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

It has to do with where the color was specified! In the first graph (which is probably what we were aiming to do), we specified color "by hand" in geom_point, which told us what color all the points should be. In the second graph, we specified it as an aesthetic mapping, so it understood "blue" as a category it should map all the points to, and then used the default color scheme to actually color the points.

## Univariate plots


```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

We can look at some new aesthetics and mappings: 

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

Density plots can be used in place of histograms. 

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

From the course website: The adjust parameter in density plots is similar to the binwidth parameter in histograms, and it helps to try a few values. I set the transparency level to .4 to make sure all densities appear. You should also note the distinction between fill and color aesthetics here. You could facet by name as above but would have to ask if that makes comparisons easier or harder. Lastly, adding geom_rug() to a density plot can be a helpful way to show the raw data in addition to the density.

Histograms and densities are one way of investigating univariate distributions; boxplots are another.

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()
```

There are also violin plots!

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

Another type of plot that's good to use if you want to visually see the distributions of individual variables is the ridge plot. 

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

### Learning assessment

```{r}
ggplot(weather_df, aes(x = prcp, fill = name)) +
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
ggplot(weather_df, aes(x = prcp, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

```{r}
ggplot(weather_df, aes(x = name, y = prcp)) + geom_boxplot()
```

```{r}
ggplot(weather_df, aes(x = name, y = prcp)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

```{r}
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

# Saving plots! 

NEVER USE THE EXPORT BUTTON. ggsave keeps things reproducible!!

```{r}
ggp_weather = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```

From the website: 

Embedding plots in an R Markdown document can also take a while to get used to, because there are several things to adjust. First is the size of the figure created by R, which is controlled using two of the three chunk options fig.width, fig.height, and fig.asp. I prefer a common width and plots that are a little wider than they are tall, so I set options to fig.width = 6 and fig.asp = .6. Second is the size of the figure inserted into your document, which is controlled using out.width or out.height. I like to have a little padding around the sides of my figures, so I set out.width = "90%". I do all this by including the following in a code snippet at the outset of my R Markdown documents.

```{r, eval = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

