---
title: "Data Visualization II"
output: github_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(patchwork)

```

## Loading data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```


This results in a dataframe with columns with `r nrow(weather_df)` observations on six variables.

## Sample plot from last time, but let's change some labels.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp in deg C",
    y = "Max daily temp in deg C", 
    color = "Location",
    caption = "Max vs min daily temp from three locations; data from NOAA"
  )
```


## Let's start to look at scales! 
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp in deg C",
    y = "Max daily temp in deg C", 
    color = "Location",
    caption = "Max vs min daily temp from three locations; data from NOAA"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15C", "0C", "15C")
  ) + 
  scale_y_continuous(
    position = "right", 
    trans = "sqrt", 
    limits = c(0,30)
  )
```

Above, we have a couple options for when/where we transform the data and limiting the data we plot. We could do this with a filter or mutate function before we plot.


## What about colors? 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp in deg C",
    y = "Max daily temp in deg C", 
    color = "Location",
    caption = "Max vs min daily temp from three locations; data from NOAA"
  ) +
  scale_color_hue(h = c(100, 200))
```

You have to be careful with colors- there are different palattes available for continuous vs discrete variables, etc. We don't really like the one above because the shades of green are way too close to one another for it to be useful. 


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp in deg C",
    y = "Max daily temp in deg C", 
    color = "Location",
    caption = "Max vs min daily temp from three locations; data from NOAA"
  ) +
  viridis::scale_color_viridis(discrete = TRUE)
```

So, why viridis? It works for folks with various kinds of color blindness, and even in grayscale!

## Themes! 

The dafult ggplot theme has no border, scale to the right, grey background, etc. What if we want it to be different?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp in deg C",
    y = "Max daily temp in deg C", 
    color = "Location",
    caption = "Max vs min daily temp from three locations; data from NOAA"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "bottom")
```

Be careful with the order- the `theme_bw` is a global reset, so we want the fine-tuning of legend placement to come *after* it. 

There are some other themes too! theme_minimal() etc. 

## Going back to settings for the general Rmd document: 

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

In addition to image sizes, we can specify default settings for ggplots like colors and themes! 

## Data arguments in geom_*

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth()
```

Because we specified the color in the aesthetics of the ggplot itself, all of those aesthetics apply to each geom layer! If we only wanted them to apply to a single geom layer, we would specify them in that specific geom. 

```{r}
nyc_weather_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

hawaii_weather_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point()+
  geom_line(data = hawaii_weather_df)

```

## `patchwork`

Faceting doesn't always work- sometimes we need to display completely different graphs next to each other. We can use the patchwork library to accomplish this!! 

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name) ##no facet on rows, facet on names for columns
```

```{r}
ggp_temp_scatter = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5)

ggp_prcop_density = 
  weather_df |>
  filter(prcp > 25) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) + 
  theme(legend.position = "bottom")

ggp_temp_scatter + ggp_prcop_density ##can put these together using the patchwork library

```

```{r}
ggp_tmax_date = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")


(ggp_temp_scatter + ggp_prcop_density) / ggp_tmax_date
```


## Data manipulation

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

We can see that the name variable is a character variable. ggplot always converts these in the background to a factor variable, which it then plots in alphabetical order. But sometimes we want to reorder the factors to get them to display in a particular order. 

```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole, WA"))
  ) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Another way uses "reorder", which can order them according to some other variable value

```{r}
weather_df |> 
  mutate(
    name = fct_reorder(name, tmax) 
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin()
```


## complicated FAS plot

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

pups_df = 
  read_csv("data/FAS_pups.csv") |> 
  janitor::clean_names()

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day"
  ) |> 
  drop_na() |> 
  mutate(
    outcome = fct_reorder(outcome, pn_day)
  ) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```

## A note about code reuse 

Always want to use the inline r functions so that if you ever update the data, everything will run without issue! 