---
title: "EDA: Numeric Summaries"
output: github_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())

weather_df
```

## Initial numeric work

```{r}
weather_df |> 
  ggplot(aes(x = prcp)) +
  geom_histogram()
```

Here are the big outliers! 

```{r}
weather_df |> 
  filter(prcp > 1000)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

why is it banded? Probably because of the way that NY and HI are measuring their data! 

## Grouping

```{r}
weather_df |> 
  group_by(name)
```

Looks invisible so far! THe only hint is `# Groups:   name [3]`

```{r}
weather_df |> 
  group_by(name, month)
```

## Counting

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(n_obs = n())
```


Like a SQL count ! 

```{r}
weather_df |> 
  count(name, name = "n_obs")
```

A shortcut that skips grouping/summarizing! 

```{r}
weather_df |> 
  count(name, month) |> 
  pivot_wider(
    names_from = name,
    values_from = n
  )
```

We're deliberately *untidying* the data to make it more human readable. 

## General summaries

```{r}
weather_df |> 
  group_by(name) |> 
  summarize(
    mean_tmax = mean(tmax)
  )
```

We get some NAs. How do we fix this? 

```{r}
weather_df |> 
  group_by(name) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  )
```

OR

```{r}
weather_df |> 
  drop_na(tmax) |> 
  group_by(name) |> 
  summarize(
    mean_tmax = mean(tmax)
  )
```


We can, once again, group by multiple variables! We can also pick different summary stats. 

```{r}
weather_df |> 
  drop_na(tmax) |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax),
    sd_tmax = sd(tmax, na.rm = TRUE),
    med_tmax = median(tmax, na.rm = TRUE)
  )
```


Let's try a graphing example. 
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() + 
  geom_line()
```


Another pivot wider example! This time, with a nice table output courtesy of knitr kable. 
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |> 
  knitr::kable(digits = 2)
```

